---
title: "Final Project"
author: "Farhan Majumder, Leila Krichel, Rui Zhou, Jeffrey Nguyen"
date: "December 5, 2017"
output:
  html_document:
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
---

# Abstract

Body size is an easily measured trait that has pervading effects for an ecosystem's structure and function because it governs ecological interactions, population abundances and distributions, and trophic exchanges. Yet, it is still debated what factors drive body size relationships in predator-prey interactions. It is often assumed that predator-prey size relationships are positively correlated in such a way where predators are usually larger than their prey. Here, we argue that predator-prey body size relationship are mediated by key predictors that may be correlated with certain habitat types, and which may not consistently produce positive associations. Further, we also ask whether a predator's feeding behaviour influences size relationships. Feeding behaviours determine the amount of resources that can be exploited from an environment, and is mediated by physiological and morphological constraints that control the amount of energy that can be allocated to body size. We explore these questions using a dataset assembled by Barnes *et al.* (2008), which summarizes marine predator-prey relationships in terms of specific interactions, mass, habitat sampled, and feeding type. Overall, we hope to determine the distribution predator-prey body size relationships across marine habitats and how it varies with feeding behaviours. 


```{r,warning=FALSE, include=FALSE}


####   THE TEAM THEME 




jrc_theme <- function() {
    theme(
        plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA),
        axis.line = element_line(),
        text = element_text(size = 11, family = "Times"),
        panel.background = element_blank(),
        panel.grid.major = element_line(colour = "grey91"),
        panel.grid.minor = element_blank(),
        legend.position = "bottom")
}


```


```{r,warning=FALSE, include=FALSE}
# Packages

library(tidyverse)
library(dplyr)
library(readxl)
library(broom)
library(ggplot2)
library(lme4)
library(MuMIn)
library(knitr)
library(plotly)
```


```{r,warning=FALSE, include=FALSE}
# Dataset (Barnes *et al.*, 2008)

Pred <- read_csv("../data/predator.csv.gz",na = c("","n/a","NA"),col_types = cols(`SD PP`=col_double()))
colnames(Pred) <- gsub(" ", "_", colnames(Pred))

Pred <- Pred %>%
mutate(Specific_habitat=gsub("Coastal bay", "Coastal Bay", Specific_habitat)) %>%
mutate(Specific_habitat=gsub("shelf", "Shelf", Specific_habitat))

pred_prey_dataset <- Pred %>%
    group_by(Predator_common_name, 
             Prey_common_name, 
             Specific_habitat, 
             Type_of_feeding_interaction, 
             Predator_lifestage, 
             Mean_PP,
             Mean_annual_temp,
             Depth) #%>%


```

# Introduction

Arguably, an organism's body size is its most important trait. Because demographic and physiological rates are often correlated with body size, this may have profound effects on ecological interactions, the flow of energy across food webs, and the dynamics and distributions of populations (Brown *et al.*, 2004; Woodward *et al.*, 2005). Furthermore, size introduces feeding constraints, such that the size of the resource will influence who can consume it, which affects the consumer's feeding behaviour and selection of prey (Pape and Bonhommeau, 2015). This may ultimately create size structure in a food web (Brose *et al.*, 2006). Consequently, a simple metric such as body size may encompass complex biological information about an ecosystem's function and structure, thus warranting further investigation. 

Body size differences are particularly apparent in predator-prey interactions. One study exploring food webs in natural communities found that in 90% of cases, predators were larger than their prey (Cohen *et al.*, 1993). There are a number of hypotheses attempting to explain this relationship. Most frequently, it is attributed to the effect of temperature and resource availability. In ectotherms, body size and temperature have been shown to negatively correlate, a pattern known as the temperature-size rule (Gibert and DeLong, 2014). Importantly, temperature also affects a resource availability because it determines an environment's productivity (Yom-Tov and Greffen, 2011). With a focus on marine and terrestrial environments, Tucker and Rogers (2014) determined that the differences in size structure in mammalian food webs could be explained by productivity. They demonstrated that marine body size relationships were on average greater than those found in terrestrial ecosystems, ascribing this to the fact that resources are generally more abundant and less structurally complex in marine environments. Here, increased resource availability allowed for more energy flow across food webs to produce larger body sizes in mammals.   

Feeding constraints, determined by a predator's feeding behaviour, could affect the amount of resources it can actually exploit from the environment. Even if resources are abundant, an organism's morphological constraints may limit how much energy it can intake, thus constraining size structure and resulting in small body size differences between predators and prey (Arim *et al.* 2007). On the other hand, pelagic ecosystems have relatively low productivity yet support predator-prey relationships with high size structure. In cases where energy flows from microscopic phytoplankton to large predatory fish, large differences in body size can still be obtained depending on a predator's feeding behaviour (Brose *et al.* 2006; McGarvey *et al.* 2016). 

Predator-prey body size relationships have been generally suggested to be positively correlated, yet empirical evidence suggests that the magnitude and direction of these relationships may vary. This suggests that there are likely various factors interacting to influence body size relationships. Using a dataset compiled by Barnes *et al.* (2008), we hope to determine what underlying factors best predict predator-prey body size relationships within various marine ecosystems. In particular, we will determine the role of marine habitats and feeding behaviours in influencing the strength of this interaction. 


```{r,echo=FALSE,fig.width=7, fig.height=6,warning=FALSE,fig.cap=paste("Fig.1: *Shows the log of Predator mass vs log of Prey Mass")}

 pred_prey_dataset %>% 
     ggplot(aes(x = log(Predator_mass), y = log(Prey_mass))) +
     geom_hex(bins = 50) +
     geom_smooth(method = "glm", colour = "Red") +
    jrc_theme()


```



```{r, fig.width=8, fig.height=5, echo=FALSE,fig.cap=paste("Fig.2: *Shows the log of Predator mass vs log of Prey Mass by Specific Habitat*")}

pred_prey_dataset %>% 
    ggplot(aes(x = log(Predator_mass), y = log(Prey_mass), colour = Specific_habitat)) +
   geom_point()


```


```{r}
Pred %>% 
     group_by(Specific_habitat) %>% 
    tally()%>% 
      ggplot(aes(x=Specific_habitat, y=n, fill = Specific_habitat))+
     geom_bar(stat = "identity")+
     jrc_theme()+
    xlab("Specific Habitat")+
    ylab("Number of Samples")+
    labs(color='Specific Habitat')+
    scale_fill_discrete(name = "Specific Habitat")
   

```


```{r, fig.width=8, fig.height=6, echo=FALSE,fig.cap=paste("Fig.4")}
# Fig. 2b
Pred %>%
    ggplot(aes(x = log(Predator_mass),
               y = log(Prey_mass),
               colour = Type_of_feeding_interaction)) +
    geom_point(size = 0.01) +
    facet_wrap(~ Specific_habitat) +
    geom_smooth(method = 'glm', colour = 'black', size = 0.5) 
    
```


# Methods

Data was obtained from the “Marine predator and prey body sizes”, created by Barnes et al(2008).; joint venture generated by 17 different international institutions. As such, we have data from global locations on organisms and their environment. We focused on data on environmental variables to see which ones suggest a correlation with predator mass and prey mass. Our variables of interest were selected to be: specific habitat location and feeding behaviour.
 
Importing the dataset into R, we used the package ggplot2 to visualize the patterns and trend in the aforementioned variables. Three graphs were generated from the data. The first figure compares predator mass and prey mass. The second figure is a map graph representing the locations of all the 27 international marine locations where our data was obtained. Using the facet feature of ggplot2, the third figure visualized the affect of feeding behaviour and the specific location on the relationship of predator-prey mass ratios.
 
To determine which variables are affecting the slope of predator-prey mass relations, we generated several models. (using glm; and lmr; and mn <- talk about these later). Every model assumes that the predator mass are the independent variable, and the prey mass are the dependent variable. The first model was a control model; assuming that there was nothing affecting the interaction of predator-prey mass. The second model incorporates the effect of specific habitat on the intercept and slope of the predator-prey mass ratio. Fourth model incorporates the effect of feeding behaviour on the intercept and slope of the predator-prey mass ratio.  The third model incorporates the additive effect of both specific habitat and feeding behaviour. The sixth model incorporates the mixed effect of habitat on feeding behaviour in the predator-prey mass interaction. The seventh model incorporates the mixed effect of feeding behaviour in the mass interaction. The last model incorporates for “everything” (<- need more details). 


# Results

The first graph shows a positive correlation between the log of predator mass and the log of prey mass; visualized via a linear regression line. As the dataset indicated, prey mass was determined off predator mass; ergo prey mass is the dependent variable.
 
Second graph is a map-graph representing the locations of the sample locations (MORE TO COME)
 
Third graph is the one that details how feeding behaviour and specific habitat affect the interaction of predator-prey mass. We can generalize the feeding behaviours of the predators into two categories: specialist feeders (the predators categorized as piscivorous, insectivorous, and herbivorous) and generalist feeders (categorized only has predacious). All the graphs under the specialist feeders showed a mild to strong positive correlation between predator and prey mass.
*In the piscivorous column, of the ten total plots, 8 of them show a positive correlation, 2 of them show neutral correlation
*In the planktivourous column, of the six plots that have a linear regression, 5 of them are positive in correlation, the last one shows a neutral correlation.
On the other hand, within the graphs under generalist feeders, there were a variety of trends; some showed positive correlations, but also some interactions were neutral and even negative in correlation.
*There are 12 plots in the column under predacious. 6 of the plots show an indication of positive trend, 5 show a neutral correlation, and one show a negative correlation. 



### Models


```{r,warning=FALSE, include=FALSE}

filtered_dataset <- pred_prey_dataset %>% 
    filter(Type_of_feeding_interaction != 'insectivorous' & Type_of_feeding_interaction != 'predacious/piscivorous') %>% 
    filter(Specific_habitat != 'Coastal, SW & SE Greenland' & Specific_habitat != 'inshore' & Specific_habitat != 'Nearshore waters')


lm9 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 | Type_of_feeding_interaction) + (1 | Type_of_feeding_interaction:Specific_habitat), data = filtered_dataset)

lm10 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 + log(Predator_mass) | Type_of_feeding_interaction) + (1 + log(Predator_mass)| Type_of_feeding_interaction:Specific_habitat), data = filtered_dataset)

lm11 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 + log(Predator_mass) | Specific_habitat) + (1 + log(Predator_mass)| Specific_habitat:Type_of_feeding_interaction), data = filtered_dataset)

lm12 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 + log(Predator_mass) | Specific_habitat) + (1 + log(Predator_mass)| Specific_habitat:Type_of_feeding_interaction) + (1 + log(Predator_mass)| Type_of_feeding_interaction), data = filtered_dataset)


fitted_lm11 <- augment(lm11) 

null <- glm(log(Prey_mass) ~ log(Predator_mass),data = filtered_dataset) 
fitted_null <- augment(null) %>% 
    rename(.fitted_null = .fitted)
names(fitted_null)
null_line <- geom_line(aes(x = log.Predator_mass., y = .fitted_null))
nulllm11 <- bind_cols(fitted_lm11, fitted_null)


model.sel(lm9, lm10, lm11,null, lm12, rank = AIC)
```



```{r,warning=FALSE, echo=FALSE}

kable(model.sel(lm9, lm10, lm11,null, lm12, rank = AIC))
```

*Table.1: Results of our models*


### Graphs 

```{r, fig.width=11, fig.height=5, warning=FALSE, echo=FALSE}
    ggplot(nulllm11, aes(x = log.Predator_mass., y = .fitted)) +
    geom_line(aes(colour = Specific_habitat))+facet_grid(~ Type_of_feeding_interaction) + null_line +
    jrc_theme()
   
```



```{r,warning=FALSE, echo=FALSE}

fitdata11 <- ranef(lm11)[[1]] %>% 
    rownames_to_column()%>% as_data_frame() %>% 
    separate(rowname, into = c("habitat","feeding_type"), sep = ":") 

colnames(fitdata11)[4] <- c("slope")

colourplz <- fitdata11 %>% 
    ggplot(aes(x = feeding_type, y = fixef(lm11)[[2]] + slope,color = habitat))+
    geom_point() + 
    geom_hline(yintercept = tidy(null)[2,2])

ggplotly(colourplz)

```

# Discussion

We analyzed predator-prey body size relationships in various marine ecosystems that spanned different latitudes and depths. The magnitude and direction of the relationships seemed to be depend on habitat and its interaction with feeding behaviour (Table 1, lm11). We assumed that different habitats correlated with certain environmental variables, allowing us to use habitat as a proxy for the environmental effect on body size relationships. We took into account the feeding behaviour of predators, as morphological and physiological constraints on a predator's feeding have been shown to be important in determining what energy predator's can actually exploit from their environment (Brose *et al.*, 2006; Arim *et al.*, 2007). 
The differences in predator-prey body size relationships suggests that different environment's have different capacities to produce resources, which thus determines how much energy is allowed to flow between predators and prey. For example, we found that piscivorous feeders in the seasonal pack ice zone have a lower intercept than piscivorous feeders in the shelf (fig. 4). Assuming that piscivores require similar energy needs, this supports the prediction that habitats with low productivity hold less energy, and thus produce food webs in which predators and prey tend to be smaller (Barnes *et al.*, 2010).

However, because habitat can encompass a wide range of variables, it was not possible to parse out exactly what variables were important in predicting predator-prey size relationships. This was due to the manner in which environmental data was collected in this dataset. For example, the average temperature for each environment only included sea surface temperature, which was obtained from an extrinsic source (Barnes *et al.*, 2008). As a result, the temperature variable does not take into account changes in temperature with depth. Thus, for samples that were observed deep in the water column, sea surface temperature would not have been informative as a predictive variable. The same logic can be applied to other variables such as the average primary productivity of an environment, which was also measured at sea surface.

On the other hand, size differences in the seasonal pack ice zone and the shelf flip when the feeding behaviour of the predator is predacious (fig. 5). Here, it is observed that shelf has a much shallower slope than both the null model and the seasonal pack ice zone, suggesting food webs regulated by predacious feeders may be more size constrained, thus producing size relationships that do not differ significantly between predators and prey. This could be a function of two different patterns. First, this pattern could be driven by the temperature-size rule. This rule states that individuals in warm environments grow more quickly due to increased metabolic rates; consequently, they reach maturity earlier, but tend to be smaller (Gibert and DeLong, 2014). In this way, size structure could be smaller in warm habitats than in cold environments like the seasonal pack ice zone. Second, this pattern may also be generated by what predacious feeders consume versus what piscivorous feeders consume. Piscivores mainly consume fish, while predacious feeders eat a more 'generalized' diet of fish, plankton, and squid. It is possible that predators with more generalized diets may be less efficient in extracting energy from their resources as their adaptations are not specialized toward a certain prey. This could reflect a tradeoff: predators with generalized diets can eat a variety of prey in their environment, but do not outperform specialists in growth and survival (Berkström *et al.*, 2014). 

A third pattern is also possible but involves limitations of the dataset. It is difficult to infer any general trend about the distribution and variation of predator-prey body size relationships across marine ecosystems and across feeding behaviours due to the dataset containing multiple incidences of nonindependent data. While the dataset contained over 30,000 entries, much of it represented nonindependent data, which may have substantially decreased the sample size. For instance, out of the whole dataset, 3581 samples were from the Albacore tuna. To account for nonindependence, we used a linear mixed effects model. Statistically, then, these samples represent a single unit -- instead of having 3581 samples, there is only 1 sample with 3581 values. As a result, this dataset was limited in what it could actually communicate in terms of how predator-prey body size relationships vary across marine habitats and feeding types. In fig. 4, the predacious subgroup in the shelf habitat has a negative correlation, suggesting that as prey become smaller, predators are becoming larger. Within the shelf habitat, there are 2022 samples, but only two populations are represented: the Atlantic bluefin tuna and the longfin squid. Because we accounted for nonindependence, the trend in the shelf environment is only being driven by these two populations, which statistically, amounts to two units. Evidently, a sample size of two is not enough for making any definitive conclusions about the shelf marine ecosystem, or about predator-prey body size relationships in the marine environment in general.

# Conclusion

In short, we can conclude that the relationship between prey and predator size can be affected by habitat and the interaction between habitat and predators’ feeding type, and interpreting the relationship by general trend may lead to Simpsons’ paradox. There might be some other potential parameters to help predict the relationship, for example, surface temperature, depth, primary productivity, etc. However, due to the limitation of our data set, we can not make a further conclusion. The sample size for each subgroup is not even, which might affect our result as well. In further, if a larger and more informative dataset can be made, it will be very helpful to generate a better model.



# Software
The software we used to create this document is the statistical Analysis program R.  The packages used to create this data are:

1)  tidyverse
2)  dplyer
3)  readxl
4)  broom
5)  lme4
6)  MuMIn
7)  knitr
8)  plotly


# Literature Cited

Arim 

1) Cohen, J. E., Pimm, S. L., Yodzis, P., & Saldana, J. (1993). *Body sizes of animal predators and animal prey in food webs*. Journal of Animal Ecology, 62, 67-78. doi:10.2307/5483

2) Brose 2006

3) Brown 2004

3) C. Barnes, D. M. Bethea, R. D. Brodeur, J. Spitz, V. Ridoux, C. Pusineri, B. C. Chase, M. E. Hunsicker, F. Juanes, A. Kellermann, J. Lancaster, F. M?nard, F.-X. Bard, P. Munk, J. K. Pinnegar, F. S. Scharf, R. A. Rountree, K. I. Stergiou, C. Sassa, A. Sabates, and S. Jennings. 2008. *Predator and prey body sizes in marine food webs*. Ecology 89:881.http://www.esapubs.org/archive/ecol/E089/051/metadata.htm

Gibert and DeLong 2014

McGarvey, Dowling, Cohen 2016

Pape and Bon

4) Tucker MA, Rogers TL. 2014, *Examining predator–prey body size, trophic level and body mass across marine and terrestrial mammals*. Proc. R. Soc. B 281:20142103., http://dx.doi.org/10.1098/rspb.2014.2103

Woodward 2005

Yom-Tov and Greffen
