---
title: "Final Project"
author: "Farhan Majumder, Leila Krichel, Rui Zhou, Jeffrey Nguyen"
date: "December 5, 2017"
output:
  html_document:
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
---


gdasgagds
```{r,warning=FALSE, include=FALSE}


####   THE TEAM THEME 




jrc_theme <- function() {
    theme(
        plot.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA),
        axis.line = element_line(),
        #text = element_text(size = 11, family = "Times"),
        panel.background = element_blank(),
        panel.grid.major = element_line(colour = "grey91"),
        panel.grid.minor = element_blank(),
        legend.position = "bottom")
}


```


```{r,warning=FALSE, include=FALSE}
# Packages

library(tidyverse)
library(dplyr)
library(readxl)
library(broom)
library(ggplot2)
library(lme4)
library(MuMIn)
library(knitr)
library(plotly)
```


```{r,warning=FALSE, include=FALSE}
# Dataset (Barnes *et al.*, 2008)

Pred <- read_csv("../data/predator.csv.gz",na = c("","n/a","NA"),col_types = cols(`SD PP`=col_double()))
colnames(Pred) <- gsub(" ", "_", colnames(Pred))

Pred <- Pred %>%
mutate(Specific_habitat=gsub("Coastal bay", "Coastal Bay", Specific_habitat)) %>%
mutate(Specific_habitat=gsub("shelf", "Shelf", Specific_habitat))

pred_prey_dataset <- Pred %>%
    group_by(Predator_common_name, 
             Prey_common_name, 
             Specific_habitat, 
             Type_of_feeding_interaction, 
             Predator_lifestage, 
             Mean_PP,
             Mean_annual_temp,
             Depth) #%>%


```

# Introduction

Arguably, an organism's body size is its most important trait. Because demographic and physiological rates are often correlated with body size, this may have profound effects on ecological interactions, the flow of energy across food webs, and the dynamics and distributions of populations (Brown *et al.*, 2004; Woodward *et al.*, 2005). Furthermore, size introduces feeding constraints, such that the size of the resource will influence who can consume it, which affects the consumer's feeding behaviour and selection of prey (Pape and Bonhommeau, 2015). This may ultimately create size structure in a food web (Brose *et al.*, 2006). Consequently, a simple metric such as body size may encompass complex biological information about an ecosystem's function and structure, thus warranting further investigation. 

Body size differences are particularly apparent in predator-prey interactions. One study exploring food webs in natural communities found that in 90% of cases, predators were larger than their prey (Cohen *et al.*, 1993). There are a number of hypotheses attempting to explain this relationship. Most frequently, it is attributed to the effect of temperature and resource availability. In ectotherms, body size and temperature have been shown to negatively correlate, a pattern known as the temperature-size rule (Gibert and DeLong, 2014). Importantly, temperature also affects a resource availability because it determines an environment's productivity (Yom-Tov and Greffen, 2011). With a focus on marine and terrestrial environments, Tucker and Rogers (2014) determined that the differences in size structure in mammalian food webs could be explained by productivity. They demonstrated that marine body size relationships were on average greater than those found in terrestrial ecosystems, ascribing this to the fact that resources are generally more abundant and less structurally complex in marine environments. Here, increased resource availability allowed for more energy flow across food webs to produce larger body sizes in mammals.   

Feeding constraints, determined by a predator's feeding behaviour, could affect the amount of resources it can actually exploit from the environment. Even if resources are abundant, an organism's morphological constraints may limit how much energy it can intake, thus constraining size structure and resulting in small body size differences between predators and prey (Arim *et al.* 2007). On the other hand, pelagic ecosystems have relatively low productivity yet support predator-prey relationships with high size structure. In cases where energy flows from microscopic phytoplankton to large predatory fish, large differences in body size can still be obtained depending on a predator's feeding behaviour (Brose *et al.* 2006; McGarvey *et al.* 2016). 

Predator-prey body size relationships have been generally suggested to be positively correlated, yet empirical evidence suggests that the magnitude and direction of these relationships may vary. This suggests that there are likely various factors interacting to influence body size relationships. Using a dataset compiled by Barnes *et al.* (2008), we hope to determine what underlying factors best predict predator-prey body size relationships within various marine ecosystems. In particular, we will determine the role of marine habitats and feeding behaviours in influencing the strength of this interaction. 


```{r,echo=FALSE,fig.width=7, fig.height=6,warning=FALSE,fig.cap=paste("Fig.1: *Shows the log of Predator mass vs log of Prey Mass")}

 pred_prey_dataset %>% 
     ggplot(aes(x = log(Predator_mass), y = log(Prey_mass))) +
     geom_hex(bins = 50) +
     geom_smooth(method = "glm", colour = "Red") +
    jrc_theme()


```



```{r, fig.width=8, fig.height=5, echo=FALSE,fig.cap=paste("Fig.2: *Shows the log of Predator mass vs log of Prey Mass by Specific Habitat*")}

pred_prey_dataset %>% 
    ggplot(aes(x = log(Predator_mass), y = log(Prey_mass), colour = Specific_habitat)) +
   geom_point()


```




```{r, fig.width=8, fig.height=6, echo=FALSE,fig.cap=paste("Fig.4")}
# Fig. 2b
Pred %>%
    ggplot(aes(x = log(Predator_mass),
               y = log(Prey_mass),
               colour = Type_of_feeding_interaction)) +
    geom_point(size = 0.01) +
    facet_wrap(~ Specific_habitat) +
    geom_smooth(method = 'glm', colour = 'black', size = 0.5) 
    
```


# Methods

Data was obtained from the “Marine predator and prey body sizes”, created by Barnes et al(2008).; joint venture generated by 17 different international institutions. As such, we have data from global locations on organisms and their environment. We focused on data on environmental variables to see which ones suggest a correlation with predator mass and prey mass. Our variables of interest were selected to be: specific habitat location and feeding behaviour.
 
Importing the dataset into R, we used the package ggplot2 to visualize the patterns and trend in the aforementioned variables. Three graphs were generated from the data. The first figure compares predator mass and prey mass. The second figure is a map graph representing the locations of all the 27 international marine locations where our data was obtained. Using the facet feature of ggplot2, the third figure visualized the affect of feeding behaviour and the specific location on the relationship of predator-prey mass ratios.
 
To determine which variables are affecting the slope of predator-prey mass relations, we generated several models. (using glm; and lmr; and mn <- talk about these later). Every model assumes that the predator mass are the independent variable, and the prey mass are the dependent variable. The first model was a control model; assuming that there was nothing affecting the interaction of predator-prey mass. The second model incorporates the effect of specific habitat on the intercept and slope of the predator-prey mass ratio. Fourth model incorporates the effect of feeding behaviour on the intercept and slope of the predator-prey mass ratio.  The third model incorporates the additive effect of both specific habitat and feeding behaviour. The sixth model incorporates the mixed effect of habitat on feeding behaviour in the predator-prey mass interaction. The seventh model incorporates the mixed effect of feeding behaviour in the mass interaction. The last model incorporates for “everything” (<- need more details). 


# Results

The first graph shows a positive correlation between the log of predator mass and the log of prey mass; visualized via a linear regression line. As the dataset indicated, prey mass was determined off predator mass; ergo prey mass is the dependent variable.
 
Second graph is a map-graph representing the locations of the sample locations (MORE TO COME)
 
Third graph is the one that details how feeding behaviour and specific habitat affect the interaction of predator-prey mass. We can generalize the feeding behaviours of the predators into two categories: specialist feeders (the predators categorized as piscivorous, insectivorous, and herbivorous) and generalist feeders (categorized only has predacious). All the graphs under the specialist feeders showed a mild to strong positive correlation between predator and prey mass.
*In the piscivorous column, of the ten total plots, 8 of them show a positive correlation, 2 of them show neutral correlation
*In the planktivourous column, of the six plots that have a linear regression, 5 of them are positive in correlation, the last one shows a neutral correlation.
On the other hand, within the graphs under generalist feeders, there were a variety of trends; some showed positive correlations, but also some interactions were neutral and even negative in correlation.
*There are 12 plots in the column under predacious. 6 of the plots show an indication of positive trend, 5 show a neutral correlation, and one show a negative correlation. 



### Models


```{r,warning=FALSE, include=FALSE}

filtered_dataset <- pred_prey_dataset %>% 
    filter(Type_of_feeding_interaction != 'insectivorous' & Type_of_feeding_interaction != 'predacious/piscivorous') %>% 
    filter(Specific_habitat != 'Coastal, SW & SE Greenland' & Specific_habitat != 'inshore' & Specific_habitat != 'Nearshore waters')


lm9 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 | Type_of_feeding_interaction) + (1 | Type_of_feeding_interaction:Specific_habitat), data = filtered_dataset)

lm10 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 + log(Predator_mass) | Type_of_feeding_interaction) + (1 + log(Predator_mass)| Type_of_feeding_interaction:Specific_habitat), data = filtered_dataset)

lm11 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 + log(Predator_mass) | Specific_habitat) + (1 + log(Predator_mass)| Specific_habitat:Type_of_feeding_interaction), data = filtered_dataset)

lm12 <- lmer(log(Prey_mass) ~ log(Predator_mass) + (1 + log(Predator_mass) | Specific_habitat) + (1 + log(Predator_mass)| Specific_habitat:Type_of_feeding_interaction) + (1 + log(Predator_mass)| Type_of_feeding_interaction), data = filtered_dataset)


fitted_lm11 <- augment(lm11) 

null <- glm(log(Prey_mass) ~ log(Predator_mass),data = filtered_dataset) 
fitted_null <- augment(null) %>% 
    rename(.fitted_null = .fitted)
names(fitted_null)
null_line <- geom_line(aes(x = log.Predator_mass., y = .fitted_null))
nulllm11 <- bind_cols(fitted_lm11, fitted_null)


model.sel(lm9, lm10, lm11,null, lm12, rank = AIC)
```



```{r,warning=FALSE, echo=FALSE}

kable(model.sel(lm9, lm10, lm11,null, lm12, rank = AIC))
```

*Table.1: Results of our models*


### Graphs 

```{r, fig.width=11, fig.height=5, warning=FALSE, echo=FALSE}
    ggplot(nulllm11, aes(x = log.Predator_mass., y = .fitted)) +
    geom_line(aes(colour = Specific_habitat))+facet_grid(~ Type_of_feeding_interaction) + null_line +
    jrc_theme()
   
```



```{r,warning=FALSE, echo=FALSE}

fitdata11 <- ranef(lm11)[[1]] %>% 
    rownames_to_column()%>% as_data_frame() %>% 
    separate(rowname, into = c("habitat","feeding_type"), sep = ":") 

colnames(fitdata11)[4] <- c("slope")

colourplz <- fitdata11 %>% 
    ggplot(aes(x = feeding_type, y = fixef(lm11)[[2]] + slope,color = habitat))+
    geom_point() + 
    geom_hline(yintercept = tidy(null)[2,2])

ggplotly(colourplz)

```

# Discussion

We analyzed predator-prey body size relationships in marine ecosystems 

- biomass relationships depend on what habitat you're in 
    - different slope for same feeding type 

- dataset limited us to studying interactions functionally
-high predation could drive smaller body size


## Caveats and Limitations 

Several limitations have been identified that may be important in the interpretation of this work. First, while this dataset had over 30,000 entries, much of the data represented nonindependent records. There were multiple records for the same predator, and further, multiple records for the same species that we took into account using a linear mixed effects model. In spite of this, nonindependent observations may have decreased the sample size, thus making it unrepresentative of food webs in marine habitats. For example, out of the whole dataset, 3581 samples are of the Albacore tuna. Statistically, these samples represent a single unit -- consequently, instead of having 3581 samples, there is only 1 sample with 3581 values. As a result, this dataset was limited in what it could actually communicate in terms of how predator-prey body size relationships vary across marine habitats and feeding types. Second, due to the manner in which environmental data was obtained for each dataset, we did not directly include it in our analysis. Instead, because environmental variables correlate with certain habitats and locations, we used specific habitat as a proxy for its effect on body size relationships. It was then not possible to infer exactly what environmetal factors were important in determining predator-prey body size relationships.


Third, because this dataset represents a compilation of work done by various studies, this .. For instance,

# Conclusion





# Software
The software we used to create this document is the statistical Analysis program R.  The packages used to create this data are:

1)  tidyverse
2)  dplyer
3)  readxl
4)  broom
5)  lme4
6)  MuMIn
7)  knitr
8)  plotly


# Literature Cited

Arim 

1) Cohen, J. E., Pimm, S. L., Yodzis, P., & Saldana, J. (1993). *Body sizes of animal predators and animal prey in food webs*. Journal of Animal Ecology, 62, 67-78. doi:10.2307/5483

2) Brose 2006

3) Brown 2004

3) C. Barnes, D. M. Bethea, R. D. Brodeur, J. Spitz, V. Ridoux, C. Pusineri, B. C. Chase, M. E. Hunsicker, F. Juanes, A. Kellermann, J. Lancaster, F. M?nard, F.-X. Bard, P. Munk, J. K. Pinnegar, F. S. Scharf, R. A. Rountree, K. I. Stergiou, C. Sassa, A. Sabates, and S. Jennings. 2008. *Predator and prey body sizes in marine food webs*. Ecology 89:881.http://www.esapubs.org/archive/ecol/E089/051/metadata.htm

Gibert and DeLong 2014

McGarvey, Dowling, Cohen 2016

Pape and Bon

4) Tucker MA, Rogers TL. 2014, *Examining predator–prey body size, trophic level and body mass across marine and terrestrial mammals*. Proc. R. Soc. B 281:20142103., http://dx.doi.org/10.1098/rspb.2014.2103

Woodward 2005

Yom-Tov and Greffen
